The origin of cosmic-ray electrons in cluster outskirts

### 2012-01-30 paper.tex
Method

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\Pp}{\mathrm{P}_\mathrm{p}}
\newcommand{\Pe}{\mathrm{P}_\mathrm{e}}
\newcommand{\pp}{p_\mathrm{p}}
\newcommand{\pe}{p_\mathrm{e}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}
\newcommand{\gap}{\gamma_\mathrm{p}}
\newcommand{\gam}{\gamma}
\newcommand{\tend}{t_f}

@@

Our scheme is summarized in decending order by the following bullet points:
\begin{itemize}
\item derive CR proton distribution function from each simulation output at different time $t$
\item calculate injected CR proton population between snapshots 
\item convert injected protons to injected electrons
\item evolve each injected CR electron population to a later time $\tend$ by
  accounting for losses (Coulomb, inverse Compton, and adiabatic)
\item add up the CR electron distribution for all SPH particles at time $\tend$
\item re-accelerate CR electron distribution function at time $\tend$
  using typical values for a merging shock in cluster outskirts
\item calculate the synchrotron radio emission from radio relic
\end{itemize}

\subsection{CR protons}
The CR protons in the simulations are represented by a one-dimensional
isotropic distribution function \footnote{The three dimensional
  distribution function is given by $4\pi \int f(\gam)\, \gam^2\, \dd \gam$.}
given by
\begin{equation}
  \label{eq:f_p}
  f_\p(\gap) = \frac{\dd^2 N_\p}{\dd \gap\,\dd V} = 
  C\, \gap^{-\alpha}\,\theta(\gap-q)\,, 
\end{equation}
and $\gap \approx P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$. Here $q$ is the mass normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization of the distribution function in units of density.

To improve the numerical efficiency of the simulations the CR proton
distribution function is parametrized in terms of adiabatic invariant
momentum cutoff $q_0$ and the adiabatic invariant Lagrangian amplitude
of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant
quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}

@@

The CR electron distribution function can not be derived directly
from the CR protons because of the different timescales of
losses. Instead we first need to derive the injected CR proton
distribution function between each snapshot.  The injected distribution
function is calculated as the change in CR normalization between a
snapshot at time $t$ and an earlier time $t-\Delta t$:
\begin{eqnarray}
  \label{eq:finj}
f_\rmn{inj,p}(t,\gap) &=& \Delta C(t)\,\frac{\rho}{m_\p}\,\rmn{X(\rho)} \,
\gap^{-\alpha}\qquad\rmn{where}\\
  \Delta  C(t) &=& C(t) - C(t-\Delta t)\,
\end{eqnarray}
and the adiabatic cooling term
$\rmn{X}(\rho)\equiv\left(\rho/\rho_0\right)^{-1/3}$. The time between
snapshots is denoted by $\Delta t$ and is shown in
Table~\ref{tab:cluster_sample} for each simulated cluster. Note that
for most clusters we can not resolve the cooling timescale of CR
electrons, and hence we over predict the abundance of electrons for
low and high momenta (see Figure~\ref{fig:e_spec_tests} for more
details) for some clusters. Although, in the region between $\gamma ~
(10-10^3)$ is small enough to resolve the cooling processes.

\subsection{CR electrons}
The main difference between the CR electrons and CR protons comes from
the less efficent acceleration of the electrons due to the smaller
gyroradius, and typically the larger losses. We derive the CR electron
distribution function by renormalizing the injected CR proton spectrum to
account for the different acceleration efficiencies as well as the
shift in momentum due to the factor $\sim 2000$ difference in mass. In
addition we model the Coulomb and radiative losses of the CR
electrons.

The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(\gae,t) &\equiv&  
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gap}(\gae,t) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gae}(\gae,t)\,Y \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd \gap}(\gap,t)
\left(\frac{\gae}{\gap}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}Y = 
\Delta C(t)\,\gae^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}Y\,.
 \nonumber\\
&& \\
\rmn{where}&& Y=\frac{m_\p}{m_\e}
\end{eqnarray}
We use that maximal 50\% of the energy available in a shock is
injected into CR protons ($\eta_\rmn{max,p}$) and a factor 10 smaller
efficiency of 5\% for the CR electrons ($\eta_\rmn{max,e}$). Note that
for readability we now drop the electron index $\e$ on the lorentz
factor $\gae$ for the electrons.

The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by instantaneously injecting a power-law of electrons at time $t_i$
and evolving it to a later time $t_f$ (for further details see
\citep{1999ApJ...520..529S}). The loss of energy for each particle is
described by
\begin{equation}
\label{eq:d_cool}
\frac{\dd \gam}{\dd t} = - b(\gam,t)\,,
\end{equation}
where the loss function $b(\gam,t)$ is dominated by Coulomb and IC
losses for the energies and gas densities relevant in this work. The
Coulomb losses are given by
\begin{eqnarray}
  \label{eq:b_C}
  b_\rmn{C}(\gam) &=& b_\rmn{C} \gam^2 
  = \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
  \left[\ln\left(\frac{m_\e c^2 \sqrt{\gam-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gam}\right)+\frac{1}{2}+
    \left.\left(\frac{\gam-1}{4\gam}\right)^2\right]\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons. Since
the Coulomb losses for relativistic electrons is almost independent of
energy (logarithmic), we approximate that $b_\rmn{C}(\gam,t)\approx
b_\rmn{C}(t)$ \footnote{We use a constant energy $\gamma=10^2$ to
  model the very weak dependence of energy, and note that the
  particular value of $\gamma$ does not matter as long as
  $\gamma\gg1$.}. We can now derive the shift in energy $\gam_i$ at
time $t_i$ to energy $\gam_f$ at time $t_f$ due to Coulomb cooling by
integrating Eqn.~(\ref{eq:d_cool}):
\begin{eqnarray}
  \label{eq:gamma_low}
  \gam_\low \equiv\int_{t_i}^{t_f}\dd t \, b_\rmn{C}(t) = 
  -\int_{\gam_i}^{\gam_f}\dd \gam' = -(\gam_f-\gam_i)\,.
\end{eqnarray}
Beacuse the snapshots are descrete in time, we approximate the low
momentum cut-off between time $t_i$ and the later time $t_f$ by
\begin{eqnarray}
  \label{eq:gamma_low_sum}
  \gam_\low \approx \sum_{j=i}^{f} \gam_\low(\Delta t_j) \approx
  \sum_{j=i}^{f} \Delta t_j \left[b_\rmn{C}(t_{j-1}) + b_\rmn{C}(t_j)\right]/2\,,
\end{eqnarray}
where $j$ denotes the summation over all snapshots between time $t_i$
and the later time $t_f$. Here we have approximated the low momentum
cut-off between two snapshots by the mean given by $\gam_\low(\Delta
t_j) = \Delta t_j \left[b_\rmn{C}(t_{j-1})+ b_\rmn{C}(t_j)\right]/2$,
where $\Delta t_j = t_j - t_{j-1}$.

We now continue with the IC losses that are given by
\begin{equation}
  \label{eq:b_IC}
  b_\IC(\gam,z) = b_\IC \gam^2 (1+z)^4
  = \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gam^2\,.
\end{equation}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB at redshift
$z=0$. Similarly to the Coulomb cooling we derive the evolution of energy 
of a particle subject to IC losses through
\begin{equation}
  \label{eq:b_IC_evolu}
\frac{\dd \gam}{\gam^2} = -b_\IC (1+z)^4\dd t\,.
\end{equation}
After a time $(t_f-t_i)$ has elapsed, all the electrons with
energies above $\gam_\ma$ have thermalized, where we derive $\gam_\ma$
through integrating Eqn.~(\ref{eq:b_IC_evolu}) from injected time
$z_i=z(t_i)$ to a later time $z_f=z(t_f)$:
\begin{equation}
  \label{eq:b_max}
   \frac{1}{\gam_\ma} \equiv \frac{1}{\gam_f} - \frac{1}{\gam_i} = 
   \frac{b_\IC}{H_0}\left[\Lambda(z_f)-\Lambda(z_i)\right]\,.
\end{equation}
Here $\Lambda(z) \approx z + 1.23\,z^2 + 0.50\,z^3 - 0.14\,z^4 -
0.04\,z^5$ in a Lambda CDM universe and
$H_0=100\,h\,\rmn{km}\,\rmn{s}^{-1}\,\rmn{Mpc}^{-1}$ is the Hubble constant.

Given an initial energy $\gam_i$ of an electron at time $t_i$,
Eqn.~(\ref{eq:d_cool}) can be integrated to give the value of $\gam_f$
at a later time $t_f$. The differential population density for
relativistic electrons is then given by
\begin{eqnarray}
\label{eq:f_evolv}
f_\rmn{inj,e}(\gam_f,t_f,t_i) &=& f_\rmn{inj,e}(\gam_i,t_i)
\left.\frac{\partial\gam_i}{\partial\gam_f}\right|_{t_f}\,,\quad \rmn{where} \\
 f_\rmn{inj,e}(\gam_i,t_i) &=& 
f_\rmn{inj,e}(\gam_f-\Delta\gam_\IC-\Delta\gam_\rmn{C},t_i) \nonumber\\
&=& f_\rmn{inj,e}(\gam_f+\frac{\gam_f^2}{\gam_\ma-\gam_f}+\gam_\low,t_i)\,,\quad \rmn{and} \\
\frac{\partial\gam_i}{\partial\gam_f} &=& 
\frac{\gam_\ma^2}{\left(\gam_\ma-\gam_f\right)^2}\,.
\end{eqnarray}
Here we have used that $\Delta\gam_\IC=-\gam_\low$ and
$\Delta\gam_\rmn{C}=\frac{-\gam_f^2}{\gam_\ma-\gam_f}$. The total electron
spectrum is derived from the sum of all individually cooled injected
spectra, starting from the time of injection $t_i$ until a later time
$t_f$,
\begin{eqnarray}
\label{eq:f_sum}
f_\rmn{tot,e}(\gam_f,t_f) = 
\left(\frac{\tilde{\rho}(t_f)}{\rho_0}\right)^{-1/3}\sum_j f_\rmn{inj,e}(\gam_f,t_f,t_j)\,.
\end{eqnarray}
Here the comoving gas density factor $\tilde{\rho}=\rho\,(z+1)^3$ takes care of the
adiabatic gains and losses, where $\rho_0=1$ is a reference comoving density in
GADGET. \del{Also note that the electrons injected at time $t=t_i$ have not
had the time to cool yet.}

>>>>

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\Pp}{\mathrm{P}_\mathrm{p}}
\newcommand{\Pe}{\mathrm{P}_\mathrm{e}}
\newcommand{\pp}{p_\mathrm{p}}
\newcommand{\pe}{p_\mathrm{e}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}
\newcommand{\gap}{\gamma_\mathrm{p}}
\newcommand{\gam}{\gamma}
\newcommand{\tend}{t_f}
\newcommand{\mach}{\mathcal{M}}

@@

Our scheme is summarized in descending order by the following bullet points:
\begin{itemize}
\item derive CR proton distribution function from each simulation output at different time $t$
\item calculate injected CR proton population between snapshots 
\item convert injected protons to injected electrons
\item evolve each injected CR electron population to a later time $\tend$ by
  accounting for losses (Coulomb, inverse Compton, and adiabatic)
\item add up the CR electron distribution for all SPH particles at time $\tend$
\item re-accelerate CR electron distribution function at time $\tend$
  using typical values for a merging shock in cluster outskirts
\item calculate the synchrotron radio emission from radio relic
\end{itemize}

\subsection{CR protons}
The CR protons in the simulations are represented by a one-dimensional
isotropic distribution function \footnote{The three dimensional
  distribution function is given by $4\pi \int f(\gam)\, \gam^2\, \dd \gam$.}
given by
\begin{equation}
  \label{eq:f_p}
  f_\p(\gap) = \frac{\dd^2 N_\p}{\dd \gap\,\dd V} = 
  C\, \gap^{-\alpha}\,\theta(\gap-q)\,, 
\end{equation}
and the Lorentz factor $\gap \approx P_\p/m_\p c$, where we have
normalized the proton momentum $P_\p$ with the proton mass
$m_\p$. Here $q$ is the mass normalized momentum cutoff, $\alpha$ the
spectral index, and $C$ the normalization of the distribution function
in units of density.

To improve the numerical efficiency of the simulations the CR proton
distribution function is parameterized in terms of adiabatic invariant
momentum cutoff $q_0$ and the adiabatic invariant Lagrangian amplitude
of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant
quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}

@@

The CR electron distribution function can not be derived directly from
the CR protons because of the different timescales of loss
processes. Instead we first need to derive the injected CR proton
distribution function between each snapshot.  The injected
distribution function is calculated as the change in CR normalization
between a snapshot at time $t$ and an earlier time $t-\Delta t$:
\begin{eqnarray}
  \label{eq:finj}
f_\rmn{inj,p}(t,\gap) &=& \Delta C_0(t)\,\frac{\rho}{m_\p}\,
\gap^{-\alpha}\,,\qquad\rmn{where}\\
  \Delta  C_-(t) &=& C_0(t) - C_0(t-\Delta t)\,.
\end{eqnarray}
The time between snapshots is denoted by $\Delta t$ and is shown in
Table~\ref{tab:cluster_sample} for each simulated cluster. Note that
for most clusters we can not resolve the cooling timescale of CR
electrons, and hence we over estimate the abundance of electrons for
low and high momenta for some clusters (see
Figure~\ref{fig:e_spec_tests} for more details). Although, in the
region between $\gamma \sim (10-10^3)$ is small enough to resolve the
cooling processes.

\subsection{CR electrons}
The main difference between the CR electrons and CR protons comes from
the less efficient acceleration of the electrons due to the smaller
gyroradius, and the larger energy losses. We derive the CR electron
distribution function by renormalizing the injected CR proton
distribution to account for the different acceleration efficiencies as
well as the shift in momentum due to the factor $X=m_\p/m_\e\sim 2000$
difference in mass. In addition we model the Coulomb and radiative
losses of the CR electrons.

\subsubsection{injection}
The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(\gae,t) &\equiv&  
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gae}(\gae,t) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gap}(\gae,t)\,X \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd \gap}(\gap,t)
\left(\frac{\gae}{\gap}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}X = 
\Delta C_0(t)\,\gae^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}X\,.
 \nonumber\\
&& 
\end{eqnarray}
We use that maximal 50\% of the energy available in a shock is
injected into CR protons ($\eta_\rmn{max,p}$) and a factor 10 smaller
efficiency of 5\% for the CR electrons ($\eta_\rmn{max,e}$). Note that
for readability we drop from here on the electron index $\e$ on the
Lorentz factor $\gae$ for the electrons.

\subsubsection{cooling}
The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by instantaneously injecting a power-law of electrons at time $t_i$
and evolving it to a later time $t_f$ (for further details see
\citep{1999ApJ...520..529S}). The loss of energy for each particle is
described by
\begin{equation}
\label{eq:d_cool}
\frac{\dd \gam}{\dd t} = - b(\gam,t)\,,
\end{equation}
where the loss function $b(\gam,t)$ is dominated by Coulomb and IC
losses for the energies and gas densities relevant in this work. The
Coulomb losses are given by
\begin{eqnarray}
  \label{eq:b_C}
  b_\rmn{C}(\gam) &=& b_\rmn{C} \gam^2 
  = \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
  \left[\ln\left(\frac{m_\e c^2 \sqrt{\gam-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gam}\right)+\frac{1}{2}+
    \left.\left(\frac{\gam-1}{4\gam}\right)^2\right]\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons. Since
the Coulomb losses for relativistic electrons is almost independent of
momentum (logarithmic), we approximate that $b_\rmn{C}(\gam,t)\approx
b_\rmn{C}(t)$ \footnote{We use a constant momentum $\gamma=10^2$ to
  model the very weak dependence of momentum, and note that the
  particular value of $\gamma$ does not matter as long as
  $\gamma\gg1$.}. We can now derive the shift in momentum $\gam_i$ at
time $t_i$ to momentum $\gam_f$ at time $t_f$ due to Coulomb cooling by
integrating Eqn.~(\ref{eq:d_cool}):
\begin{eqnarray}
  \label{eq:gamma_low}
  \gam_\low \equiv\int_{t_i}^{t_f}\dd t \, b_\rmn{C}(t) = 
  -\int_{\gam_i}^{\gam_f}\dd \gam' = -(\gam_f-\gam_i)\,.
\end{eqnarray}
Because the snapshots are discrete in time, we approximate the low
momentum cut-off between time $t_i$ and the later time $t_f$ by
\begin{eqnarray}
  \label{eq:gamma_low_sum}
  \gam_\low \approx \sum_{j=i+1}^{f} \gam_\low(\Delta t_j) \approx
  \sum_{j=i+1}^{f} \Delta t_j \left[b_\rmn{C}(t_{j-1}) + b_\rmn{C}(t_j)\right]/2\,,
\end{eqnarray}
where $j$ denotes the summation over all snapshots between time $t_i$
and the later time $t_f$. Here we have approximated the low momentum
cut-off between two snapshots by the mean given by $\gam_\low(\Delta
t_j) = \Delta t_j \left[b_\rmn{C}(t_{j-1})+ b_\rmn{C}(t_j)\right]/2$,
where $\Delta t_j = t_j - t_{j-1}$.

We now continue with the IC losses that are given by
\begin{equation}
  \label{eq:b_IC}
  b_\IC(\gam,z) = b_\IC \gam^2 (1+z)^4
  = \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gam^2\,.
\end{equation}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB at redshift
$z=0$. Similarly to the Coulomb cooling we derive the evolution of energy 
of a particle subject to IC losses through
\begin{equation}
  \label{eq:b_IC_evolu}
\frac{\dd \gam}{\gam^2} = -b_\IC (1+z)^4\dd t\,.
\end{equation}
After a time $(t_f-t_i)$ has elapsed, all the electrons with momentum
above $\gam_\ma$ have thermalized, where we derive $\gam_\ma$ through
integrating Eqn.~(\ref{eq:b_IC_evolu}) from the time $z_i=z(t_i)$
where the electrons are injected to a later time $z_f=z(t_f)$:
\begin{equation}
  \label{eq:b_max}
   \frac{1}{\gam_\ma} \equiv \frac{1}{\gam_f} - \frac{1}{\gam_i} = 
   \frac{b_\IC}{H_0}\left[\Lambda(z_f)-\Lambda(z_i)\right]\,.
\end{equation}
Here $\Lambda(z) \approx z + 1.23\,z^2 + 0.50\,z^3 - 0.14\,z^4 -
0.04\,z^5$ in a Lambda CDM universe and
$H_0=100\,h\,\rmn{km}\,\rmn{s}^{-1}\,\rmn{Mpc}^{-1}$ is the Hubble constant.

Given an initial energy $\gam_i$ of an electron at time $t_i$,
Eqn.~(\ref{eq:d_cool}) can be integrated to give the value of $\gam_f$
at a later time $t_f$. The differential population density for
relativistic electrons is then given by
\begin{eqnarray}
\label{eq:f_evolv}
f_\rmn{inj,e}(\gam_f,t_f,t_i) &=& f_\rmn{inj,e}(\gam_i,t_i)
\left.\frac{\partial\gam_i}{\partial\gam_f}\right|_{t_f}\,,\quad \rmn{where} \\
 f_\rmn{inj,e}(\gam_i,t_i) &=& 
f_\rmn{inj,e}(\gam_f-\Delta\gam_\IC-\Delta\gam_\rmn{C},t_i) \nonumber\\
&=& f_\rmn{inj,e}(\gam_f+\frac{\gam_f^2}{\gam_\ma-\gam_f}+\gam_\low,t_i)\,,\quad \rmn{and} \\
\frac{\partial\gam_i}{\partial\gam_f} &=& 
\frac{\gam_\ma^2}{\left(\gam_\ma-\gam_f\right)^2}\,.
\end{eqnarray}
Here we have used that $\Delta\gam_\IC=-\gam_\low$ and
$\Delta\gam_\rmn{C}=\frac{-\gam_f^2}{\gam_\ma-\gam_f}$. The total electron
spectrum is derived from the sum of all individually cooled injected
spectra, starting from the time of injection $t_i$ until a later time
$t_f$,
\begin{eqnarray}
\label{eq:f_sum}
f_\rmn{tot,e}(\gam_f,t_f) = 
\left(\frac{\tilde{\rho}(t_f)}{\rho_0}\right)^{-1/3}\sum_j f_\rmn{inj,e}(\gam_f,t_f,t_j)\,.
\end{eqnarray}
Here the comoving gas density factor $\tilde{\rho}=\rho\,(z+1)^3$
takes care of the adiabatic gains and losses (see Eqns.\ref{eq:C0} and
\ref{eq:q0}), where $\rho_0=1$ is a reference comoving density in
GADGET. \del{Also note that the electrons injected at time $t=t_i$
  have not had the time to cool yet.}
